### 2024-04-02 outputs/paper/paper.qmd
updates to discussion

---
title: "Relationship Between Stock Prices and Common Financial Metrics"
subtitle: "Analysis of the Relationship Between Stock Price and EPS & Dividends of the Largest 50 North American Companies over 10 Years"
author: Diana Liu
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(modelsummary)
library(bayesplot)
library(broom)
library(gridExtra)
library(rstanarm)
library(knitr)
#
library(gt)
library(grDevices)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data
Public companies whose stock are traded on exchanges, must disclose certain financial metrics in their quarterly and yearly financial statements. Investors then use these metrics to help inform their decision when selecting companies to invest in. Some of the most important metrics investors look at are Earnings Per Share, Net Income, and Dividends [@CiteShakespeare]. Investments take the form of buying shares or stocks which allows the investor to own a fraction of the company.

The data used for analysis is obtained through Walton Research Data Services (WRDS) by the University of Pennsylvania. WRDS provides access to Compustat, a database of financial, statistical, and market information on global companies since 1962. Our data comes from Compustat's North America Fundamentals Annual database, which contains both financial data from all public North American companies, collected by Compustat from each company's annual financial statements or from stock exchanges (@fig-Analysis-Data-Sample). From this data set, we extract the price, EPS, net income, dividends, TIC, and Year of the top 50 largest companies by market capitalization, a measure of the size of the company in terms of total value, for the past 10 years (2013-2023). This Data is cleaned and analysed in @CiteR with assistance from @CiteTidy, @CiteArrow, @CiteModelsummary, @CiteBayesplot, @CiteBroom, @CiteGridextra, @CiteRstanarm, @CiteKnitr

Similar data sets exist, but WRDS was selected because it hosts the most robust collection in terms of variables and its infrastructure allows users to easily query specific items.

```{r}
#| label: fig-Analysis-Data-Sample
#| fig-cap: "Data from Compustat North America Fundamentals Annual. EPS, Dividends, and Net Income are financial statement data while Price are stock exchange data, and Year and Tic can be found on both."
#| echo: false

analysis_data <- read_parquet(here::here("outputs/data/analysis_data.parquet"))
kable(head(analysis_data))
```

## Financial Statement Data
Net income, EPS, and dividends are items that are found on financial statements. Specifically, EPS and net income can be found on the income statement while dividends can be found on the statement of retained earnings. This data is likely unbiased and free from error as financial statements are required by law to be audited (verified by an independent third party).

Net income is used to measure profitability or how much income the company keeps after expenses are paid. Higher net income means that the company is profitable by either earning more revenues or reducing expenses. Net income is expected to be positively correlated with price as investors want to invest in companies that are more profitable. Net income can be negative when the company has lost money. In our data set, it is measured in millions USD (x10^6). Net income is normally distributed with median of 7120M USD and right skew, mean of 11752 (@fig-distributions). Apple (AAPL) had the highest net income in the data set of 99803 achieved in 2022, and Berkshire Hathaway (BRK.B) had the lowest of -22819 in the same year (@tbl-summary-statistics).

```{r}
#| label: fig-EPS-boxplot
#| fig-cap: "Boxplots of EPS grouped by dividends"
#| echo: false
EPS_boxplot <-
  analysis_data |> 
  ggplot(aes(x = Paid_Dividend, y = EPS)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Dividends Paid",
       y = "EPS")

ylim1 = boxplot.stats(analysis_data$EPS)$stats[c(1, 5)]
EPS_boxplot = EPS_boxplot + coord_cartesian(ylim = ylim1*1.05)
EPS_boxplot

```

EPS is a commonly used measure of a given company's value in USD; it is calculated net income divided by the number of shares. It is normally distributed and right skewed, with median 4.56 USD and mean 6.2 (@fig-distributions). Google (GOOGL) had the highest EPS of 112.2 in 2021, and Tesla (TSLA) had the lowest of -11.83 in 2017, meaning for each share Tesla lost 11.83 USD (@tbl-summary-statistics).

Dividends are distributions of a company's income to the owners of its stock in USD. Some companies choose not to pay dividends if they can use their income to reinvest into valuable projects. For our analysis we created a binary variable for whether or not a company paid dividends, with any amount greater than zero dollars being Yes. Approximately 78% of companies in our data set paid dividends, the largest of which was 25999 USD paid by Caterpillar (CAT) in 2023.

EPS of companies that did not pay dividends has a larger interquartile range than EPS of companies that paid dividends (@fig-EPS-boxplot). This means that the observations for companies that paid dividends are more clustered together, likely resulting in smaller standard deviation. The mean and median EPS of companies that paid dividends is also higher, suggesting that paying dividends is correlated with having higher EPS (@fig-EPS-boxplot), this makes sense as companies that are profitable enough to have high EPS also have more income to pay out dividends.

```{r}
#| include: false

count(analysis_data)
analysis_data %>% filter(Paid_Dividend == "Yes") %>% nrow()

analysis_data %>% filter(Paid_Dividend == "Yes") %>% nrow()/count(analysis_data)
```

## Stock Exchange Data
The price is the amount that each share costs to buy, it is determined by supply and demand for a company. For example, if investors believe a company is profitable, they will buy shares with the goal of eventually sharing in said profit, increasing demand and driving up share price. Share price is normally distributed with median of 113.91 USD and mean of 6.2. The most expensive stock was Amazon (AMZN) which was worth 3334.34 USD in 2021, and the cheapest stock was Advanced Micro Devices, Inc. (AMD) which was worth 2.67 in 2014.

Tic refers to the ticker symbol of each stock. It is a unique string of one to five characters used to denote a particular company's stock so that investors know whose stock are being traded on international exchanges across different languages. Some tick symbols are similar to the name of the company like AAPL for Apple and GOOGL for Google, but this is not always the case, for example American Express is AXP and Coca-cola is KO.

```{r}
#| echo: false
#| eval: true
#| label: tbl-summary-statistics
#| tbl-cap: "Summary statistics of Price, EPS, Net Income, and Dividends of the 50 largest North American Companies by market cap"
#| warning: false

earnings_per_share <- unclass(summary(analysis_data$EPS, digits = 3))
net_income <- unclass(summary(analysis_data$Net_Income, digits = 7))
dividends <- unclass(summary(analysis_data$Dividends, digits = 6))
price <- unclass(summary(analysis_data$Price, digits = 5))
summary <- data.frame(earnings_per_share, net_income, dividends, price, check.names = FALSE, stringsAsFactors = FALSE)

summary_stats <- c("Min", "1st Quartile", "Median", "Mean", "3rd Quartile", "Max")
summary <-
  cols_add(gt(summary), summary_stats) |>
  cols_move_to_start(summary_stats)

summary
```

```{r}
#| label: fig-distributions
#| fig-cap: "Price, EPS, Net_Income, and Dividends are all normally distributed with right skew"
#| echo: false
#| warning: false
#| message: false

distribution_price <-
  ggplot(analysis_data, aes(x = Price)) +
  geom_histogram() +
  theme_minimal()

distribution_EPS <-
  ggplot(analysis_data, aes(x = EPS)) +
  geom_histogram() +
  theme_minimal()

distribution_Dividends <-
  ggplot(analysis_data, aes(x = Dividends)) +
  geom_histogram() +
  theme_minimal()

distribution_Net_Income <-
  ggplot(analysis_data, aes(x = Net_Income)) +
  geom_histogram() +
  theme_minimal()

grid.arrange(distribution_price, distribution_EPS, distribution_Net_Income, distribution_Dividends, ncol = 2)
```

# Model

The goal of our model is to determine the relationship between EPS and stock price first with a simple linear model. Then we add a binary variable of whether the company paid dividends or not to determine the effect dividends have on this relationship. Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up
In the first model, we define $y_i$ as the stock price, $X_1$ as earnings per share, a continuous variable. $\beta_0$ is the intercept term, and $\beta_1$ is the coefficient for change in the stock price per dollar increase in EPS. $\epsilon$ denotes the noise or deviations.

\begin{align}
&Simple\ Model \\
&Y_i = \beta_0 + \beta_1X_1 + \epsilon \\
\end{align}

In the second model, we add $X_2$ as a binary variable where yes signifies that a company paid dividends, and $\beta_2$ is the coefficient for change in stock price if dividends were paid.

\begin{align} 
&Multivariable\ Model \\
&Y_i = \beta_0 + \beta_1X_1 + \beta_2X_2 + \epsilon \\
\end{align}

### Model justification
We expect a positive linear relationship between EPS and stock price. This is because a higher EPS indicates greater value and investors should be willing to pay more for a company's shares if they think the company has higher profits relative to its share price. If a company has high EPS, they have greater amount of profit after expenses have been paid. Investors want to own profitable companies, so demand for this company's stock increases, driving up price as the number of stocks that are available to be bought remains the same. The opposite is true for unprofitable companies, if EPS is low or negative, investors may choose to sell off their shares even at a loss in order to prevent further losses should the stock price keep falling (@fig-bestfit).

One might expect dividends to be positively correlated with stock price because investors desire an additional payment, but the opposite is true. This is because dividends paid out to shareholders is value that is taken out of the company, causing price, a reflection of the firm's value, to decrease (@fig-bestfit).

We expect the effects of EPS and dividends to interact with each other, creating two potential outcomes for change in price. Companies that did not pay dividends are expected to have a larger increase in price for every additional dollar of EPS than companies that paid dividends.

# Results
```{r}
#| include: false
#| eval: true
#| warning: false

model_1 <-
  lm(
    Price ~ EPS,
    data = analysis_data
  )

saveRDS(
  model_1,
  file = "model_1.rds"
)
```

```{r}
#| include: false
#| eval: true
#| warning: false
model_rstanarm_2 <-
  stan_glm(
    formula = Price ~ EPS + Paid_Dividend,
    data = analysis_data,
    family = gaussian(),
    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
    prior_intercept = normal(0, 2.5, autoscale = TRUE),
    prior_aux = exponential(rate = 1, autoscale = TRUE),
    seed = 12345
  )

saveRDS(
  model_rstanarm_2,
  file = "model_rstanarm_2.rds"
)
```

```{r}
#| include: false
#| eval: true

summary(model_1) 
summary(model_rstanarm_2)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-bestfit
#| fig-cap: "Relationship between variables EPS and stock price"

base_eps_plot <-
  analysis_data |> 
  ggplot(aes(x = EPS, y = Price)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(x = "EPS",
       y = "Stock Price")

base_eps_plot +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = "y ~ x"
  )
```

In model 1, only the relationship between the dependent variable price and predictor variable EPS is analyzed. In the first column of @tbl-modelresults, we have a intercept of 27.25 and coefficient of 26.59 with S.D. 1.05. The coefficient term $\beta_1$ which shows the increase in price for every dollar increase in EPS can be observed from @fig-bestfit as the dashed line with standard error as the shaded area around the dashed line. Positive slope and coefficient show a positive relationship between EPS and stock price. While most observations follow the linear relationship, there is a large cluster around 0.

```{r}
#| echo: false
#| eval: true
#| label: fig-boxplot
#| fig-cap: "Boxplots of stock price grouped by whether or not dividends were paid."
#| warning: false

dividends_plot <-
  analysis_data |> 
  ggplot(aes(x = Paid_Dividend, y = Price)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Dividends Paid",
       y = "Stock Price")

ylim1 = boxplot.stats(analysis_data$Price)$stats[c(1, 5)]
dividends_plot = dividends_plot + coord_cartesian(ylim = ylim1*1.05)
dividends_plot
```

```{r}
#| echo: false
#| eval: true
#| label: fig-model-2
#| fig-cap: "Linear regression "
#| warning: false

base <-
  analysis_data |>
  ggplot(aes(x = EPS, y = Price)) +
  labs(
    x = "EPS",
    y = "Price"
  ) +
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

base +
  geom_point(aes(color = Paid_Dividend)) +
  geom_smooth(
    aes(color = Paid_Dividend),
    method = "lm",
    linetype = "dashed",
    formula = "y ~ x"
  ) +
  labs(color = "Paid Dividends")
```

In model 2, we introduce dividends as an additional predictor variable. Stocks that paid dividends have median price around 100, significantly lower than stocks that paid no dividends with median price around 175 (@fig-boxplot). The second column of @tbl-modelresults shows that the new intercept is 172.05 USD, and that the coefficient of paying dividends $\beta_2$ is -177.85. EPS of companies that paid dividends has intercept of 172.05 and coefficient of 25.59 (@tbl-modelresults). The new standard deviation for EPS is 1.0 and 21.6 for dividends.

The effects on price from EPS and dividends can be further observed in @fig-model-2 where the observations and line of best fit are colored blue to indicate that dividends were paid and red if not. EPS without dividends (red line) has very similar slope and coefficient as @fig-bestfit, and EPS with dividends (blue line) is below.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Effect of earnings per share and dividends on stock price."
#| warning: false

modelsummary(
  list(
    "EPS Only" = model_1,
    "With Dividends" = model_rstanarm_2
  ),
  fmt = 2
)
```

# Discussion
## Companies with Higher Earnings per Share are More Attractive to Investors
From model 1, we obtain a coefficient of 21.79 and standard deviation of 1.1 for EPS, indicating that for every additional dollar increase in EPS, stock price increases by 21.79 USD with average deviation of 1.1 (@tbl-modelresults). This confirms that there is a positive linear relationship between EPS and stock price. When a company is profitable relative to the amount of shares they have, each shareholder is entitle to a larger portion of their profit, making the company a more attractive investment than competitors that have lower EPS. This demand for shares of the company then drives up stock prices. Alternatively, a lower EPS indicating lower profitability or losses can cause existing shareholders to sell their shares, creating downward pressure on price.

This conclusion has several implications for investors and management of companies alike. For investors, it may be useful to use the causal relationship between EPS and stock price to predict future stock prices. An investor can use available net income information and number of shares from any company to calculate EPS before the information is released to the public. If the estimated EPS is high, the investor can buy up shares at a lower price before the increase in demand and sell those shares once demand increases to make a profit.

Managers of a company who have an incentive for their company to be profitable in the long term would want their share prices to exhibit stable growth over time instead of the short term fluctuations caused by changing demand. If managers anticipates that one year's EPS will be low, the company can use excess income to buy up shares once the price decreases. This alleviates some of the downward pressure on price and creates more stability. The company can then either hold onto the purchased share until a subsequent year when the price rebounds to sell them, or eliminate those shares from the market. Fewer shares means that the denominator of the EPS equation is smaller, causing subsequent EPS to be higher in the future.

## Paying Dividends Causes a Decrease in Stock Price
From @fig-boxplot there is a negative relationship between paying dividends and stock price. When a company chooses to pay dividends, on average their stock price is expected to decrease by about 177.85 USD with standard deviation of 21.6 USD (@tbl-modelresults). This is consistent with contemporary literature on dividend policy which suggests that paying dividends causes the stock price to drop by the amount of the dividend per share plus the opportunity cost of lost potential profits the dividend could have been used to generate (@CiteNarinderp). Dividends paid out represents value that is extracted from the company and not used to invest into profitable projects. All things equal, this indicates that the company is expected to grow less than competitors who did not pay dividends and reinvested, making the stock less attractive to potential investors who care about growth.

EPS coefficient of companies that paid dividends is 25.59 and new standard deviation for EPS is 1.0 (@tbl-modelresults). This is smaller than that of model one where dividends were not considered. This means that the amount price is expected to increase per dollar EPS increases is larger, likely due to the fact that if a firm is profitable enough to pay dividends, it should also have high EPS.

Managers are faced with the decision of what to do with extra profit. If paid out in dividends, the stock price will fall but existing shareholders, especially large ones with 10% or more equity in the company, would want the dividend income as compensation for their investment. At the same time, paying out dividends makes the stock less attractive to future investors. By knowing the coefficient of paying dividends and coefficient of EPS increases, managers are better able to balance these competing priorities and make better decisions in regards to long-term profitability and growth of their firms.

## Weaknesses and next steps
A weakness in our model is the distribution of the data where they are clustered around zero, this may indicate that a linear model is not appropriate for our observations and the conclusions between EPS, dividends and price cannot be drawn. While the residuals follow normal distribution, they are clustered together instead of randomly spaced apart, indicating that they contain some additional structure that was not captured by our analysis (@fig-residuals). The posterior prediction check also supports this as the actual data is far larger kurtosis than what the model predicted, meaning that the predictive power of the model is low (@fig-ppcheckandposteriorvsprior-1). The credibility intervals for the coefficient of paying dividends is also quite large, with probability normally distributed between about -275 and -125 USD, suggesting a relatively high level of uncertainty (@fig-credible-intervals). Practically, the model worked as intended with no signs of problems from the trace plot or r-hat values (@fig-trace-rhat). Further work is needed to confirm the positive linear relationship between stock price and EPS.

The conclusions drawn from our analysis only applies to our data set of large, North American companies from 2013-2023. More analysis with diverse data sets in terms of need to be done to determine if the same conclusions are applicable to companies of different sizes or different geographic locations. The data is also historic, meaning that the relationships described may not occur in the future. Training the model to have predictive power may provide additional useful information.

Lastly, while financial metrics are important to management and investor decision making, there are many quantitative factors like the strategic direction of a firm and its competitive advantage within the industry should also be considered. Future analysis should aim to quantify these categorical factors and determine their relationship with stock price as well as how they interact with financial metrics.

\newpage

\appendix

# Appendix

# Model details {#sec-model-details}

```{r}
#| include: false
#| eval: true

prior_summary(model_rstanarm_2)
```

## Residuals
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-residuals
#| layout-ncol: 2
#| layout-nrow: 2
#| fig-cap: "Examining how model 2 fits the data, and how it is affected by the data"
#| fig-subcap: ["Distribution of residuals", " Residuals by EPS", "Residuals by Price", "Comparing the estimated price with the actual price"]

analysis_data <-
  augment(
    model_1,
    data = analysis_data
  )

ggplot(analysis_data, aes(x = .resid)) +
  geom_histogram(binwidth = 1) +
  theme_classic() +
  labs(y = "Number of occurrences", x = "Residuals")

ggplot(analysis_data, aes(x = EPS, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
  theme_classic() +
  labs(y = "Residuals", x = "EPS")

ggplot(analysis_data, aes(x = Price, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
  theme_classic() +
  labs(y = "Residuals", x = "Stock Price")

ggplot(analysis_data, aes(x = Price, y = .fitted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_classic() +
  labs(y = "Estimated Stock Price", x = "Actual Stock Price")
```

## Posterior predictive check
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how model 2 fits the data, and how it is affected by the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(model_rstanarm_2) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(model_rstanarm_2) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```
@fig-ppcheckandposteriorvsprior-1 compares actual dependent variable outcomes with distributions from the posterior simulation. We can see that the general shape and distribution are similar but actual data has far higher kurtosis than the simulations.

@fig-ppcheckandposteriorvsprior-2 compares priors with posteriors to see how much the estimates change once data are taken into account. Our priors and posteriors are quite similar,

## Credible Intervals
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-credible-intervals
#| fig-cap: "95% credibility intervals of model 2 variable coefficients"

plot(
  model_rstanarm_2,
  "areas"
)
```

@fig-credible-intervals shows the distributions of the 95% credibility intervals of coefficients of the predictor variables of model 2. The Bayesian estimation provides distributions of each coefficient. The actual value has a 95% probability of being within the interval with the distribution showing the likelihood within the interval. Ones that have a wider credibility interval are more spread out like dividends and intercept means that the true coefficient is within a larger range.

## Diagnostics
```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace-rhat
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(model_rstanarm_2, "trace")

plot(model_rstanarm_2, "rhat")
```
Markov chain Monte Carlo (MCMC) to was used in the model to obtain samples from the posterior distributions of interest. @fig-trace-rhat-1 shows bouncing horizontal lines with overlap, and @fig-trace-rhat-2 shows R-hat close to one. Both indicate that there are no issues with the algorithm.

\newpage
# References